Redirecting Resolution Proofs

### 2011-05-10 bib/prooftransform/redirecting-proofs-2/paper.tex
test -- trivial change
more work on redirection examples

\newdimen\bls
\bls=\baselineskip
\newdimen\forhave
\forhave=.8em %% TYPESETTING

@@

\newcommand\smallbot{\kern-.125em\bot\kern-.125em}
\let\B=\overline
%\newcommand\have{\mathrel{\vdash}}
%\newcommand\have{\mathrel{\;\hookrightarrow\;}}
\newcommand\have{\mathrel{\;\vartriangleright\;}}
\newcommand\hence{\mathrel{\;\blacktriangleright\;}}
%\newcommand\have{\mathrel{\,\rightharpoonup\,}}
\newcommand\havebot{\have\smallbot} %% TYPESETTING
\newcommand\hencebot{\hence\smallbot} %% TYPESETTING

@@

\paragraph{Proof trees.} Proof trees are the standard notation for natural
deduction proofs. For example, the proof graph above is captured by the
following proof tree:
%
\begin{prooftree}
\AXC{$[\B{0}]$}
\UIC{$\B{0}$}
    \AXC{$1$}
        \AXC{$2$}
    \BIC{$1 \mathrel{\land} 2$}
        \AXC{$1 \mathrel{\land} 2 \longrightarrow 4$}
    \BIC{$4$}
\BIC{$\B{0} \mathrel{\land} 4$}
    \AXC{$\kern-4em\B{0} \mathrel{\land} 4 \longrightarrow \B{6}$}
\BIC{$\B{6}$}
    \AXC{$[\B{0}]$}
    \UIC{$\B{0}$}
        \AXC{$3$}
            \AXC{$1$}
                \AXC{$2$}
            \BIC{$1 \mathrel{\land} 2$}
                \AXC{$1 \mathrel{\land} 2 \longrightarrow 4$}
            \BIC{$4$}
        \BIC{$3 \mathrel{\land} 4$}
            \AXC{$\kern-3em 3 \mathrel{\land} 4 \longrightarrow 5$}
        \BIC{$5$}
    \BIC{$\B{0} \mathrel{\land} 5$}
        \AXC{$\kern-5em\B{0} \mathrel{\land} 5 \longrightarrow \B{7}$}
    \BIC{$\B{7}$}
\BIC{$\B{6} \mathrel{\land} \B{7}$}
    \AXC{$\kern-5em\B{6} \mathrel{\land} \B{7} \longrightarrow \bot$}
\BIC{$\bot$}
\end{prooftree}
The proof graph notation is more compact, not least because it enables
sharing of subproofs. For that reason, we will prefer them to proof trees.

\paragraph{Isar proofs.} Isar proofs are a linearization of natural deduction
proofs, but unlike proof trees they enable sharing:
%
\begin{quote}
\keyw{proof}\,~\textit{neg\_clausify} \\
\hbox{}\quad \keyw{assume}\, $\B{0}$ \\
\hbox{}\quad \keyw{have}\, $4$ \keyw{\,by\,} (\textit{metis\, }$1$ $2$) \\
\hbox{}\quad \keyw{have}\, $5$ \keyw{\,by\,} (\textit{metis\, }$3$ $4$) \\
\hbox{}\quad \keyw{have}\, $\B{6}$ \keyw{\,by\,} (\textit{metis\, }$\B{0}$ $4$) \\
\hbox{}\quad \keyw{have}\, $\B{7}$ \keyw{\,by\,} (\textit{metis\, }$\B{0}$ $5$) \\
\hbox{}\quad \keyw{show} \kern.03em$\bot$ \keyw{\kern.05em by\,} (\textit{metis\, }$\B{6}$ $\B{7}$) \\ %% TYPESETTING
\keyw{qed}
\end{quote}
%
The above proof is by contradiction. A direct proof of the above in Isar would
be
%
\begin{quote}
\keyw{proof} -- \\
\hbox{}\quad \keyw{have}\, $4$ \keyw{\,by\,} (\textit{metis\, }$1$ $2$) \\
\hbox{}\quad \keyw{have}\, $5$ \keyw{\,by\,} (\textit{metis\, }$3$ $4$) \\
\hbox{}\quad \keyw{have}\, $6 \mathrel{\lor} 7$ \keyw{\,by\,} \textit{metis} \\
\hbox{}\quad $\{$\quad \keyw{assume}\, $6$ \\
\hbox{}\qquad\phantom{$\{$}\keyw{have}\, $0$ \keyw{\,by\,} (\textit{metis\, }$4$ $6$)\quad$\}$ \\
\hbox{}\quad \keyw{moreover} \\
\hbox{}\quad $\{$\quad \keyw{assume}\, $7$ \\
\hbox{}\qquad\phantom{$\{$}\keyw{have}\, $0$ \keyw{\,by\,} (\textit{metis\, }$5$ $7$)\quad$\}$ \\
\hbox{}\quad \keyw{ultimately show}\, $0$ \keyw{\,by\,} (\textit{metis\, }`$6 \mathrel{\lor} 7$') \\
\keyw{qed}
\end{quote}
We refer to the Isar tutorial \cite{nipkow-2011-isar} for more information on
the Isar syntax.

@@

We start with a simple proof by contradiction expressed as a proof tree and
using our shorthand notation:
%
\[\hbox{\begin{tabular}{l@{\qquad}l}
\raisebox{-\height}{\includegraphics[scale=0.45]{diagrams/linear-ref-graph.pdf}}
&
\qquad
\raisebox{-\height}{$\begin{aligned}[t]
\\
\B 0, 1 & \have \B 3 \\
2, \B 3 & \have \B 4 \\
1, \B 4 & \havebot
\end{aligned}$}
\end{tabular}}\]
Recall from Sect.\ \ref{sec:proof-notations} that the negated conjecture all all
the proof steps that are tainted by it are shown as negated. Next, we turn the
sequents around using contraposition at the sequent level to avoid all
negations. This gives
\begin{align*}
1, 3 & \have 0 \\
2, 4 & \have 3 \\
1 & \have 4
\end{align*}
Finally, if we reorder the sequents and introduce $\hence$ wherever possible, we
obtain a direct proof:

@@

The next two examples are lasso-shaped proofs:
%
\[\hbox{\begin{tabular}{llll}
\raisebox{-\height}{\includegraphics[scale=0.45]{diagrams/lasso-1-ref-graph.pdf}}
&
\qquad
\raisebox{-\height}{$\begin{aligned}[t]
\\
\B 0 & \have \B 1 \\
\B 0 & \have \B 2 \\
\B 1 & \have \B 3 \\
\B 2 & \have \B 4 \\
\B 3, \B 4 & \have \B 5 \\
\B 5 & \have \B 6 \\
\B 6 & \havebot
\end{aligned}$}
&
\qquad\qquad
\raisebox{-\height}{\includegraphics[scale=0.45]{diagrams/lasso-2-ref-graph.pdf}}
&
\qquad
\raisebox{-\height}{$\begin{aligned}[t]
\\
\B 0 & \have \B 1 \\
\B 1 & \have \B 2 \\
\B 2 & \have \B 3 \\
\B 2 & \have \B 4 \\
\B 3 & \have \B 5 \\
\B 4 & \have \B 6 \\
\B 5, \B 6 & \havebot
\end{aligned}$}
\end{tabular}}\]

We start with the contradiction proof on the left-hand side. Starting from
$\bot$, it is easy to turn the proof around up to the lasso cycle:
%
\begin{align*}
& \have 6 \\
& \hence 5 \\
& \hence 3 \lor 4
\end{align*}
%
When applying the contrapositive to eliminate the negations in $\B 3, \B 4 \have
\B 5$, we obtain a disjunction on the right-hand side of the sequent: $5 \have 3
\lor 4$. To continue from there, we need a case split. Then we can finish then
branches:
%
\[\left[\begin{tabular}{l|l} %% TYPESETTING
    $\begin{aligned}[t]
      [&3] \\
      \hence {} & 1 \\
      \hence {} & 0
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t] 
      [&4] \\
      \hence {} & 2 \\
      \hence {} & 0
    \end{aligned}$
  \end{tabular}\right]\]

@@

In this case, the key is to observe that it is enough if both branches prove
$2$, and from there we prove the rest. The complete proof is (without and
with $\hence$):
\[\hbox{\begin{tabular}{@{}p{.25\textwidth}@{}p{.25\textwidth}@{}}
\frm \hav 5 \lor 6 \end
\cases{\begin{tabular}{l|l}
    $\begin{aligned}[t]
      [&5] \\
      5 \have {} & 3 \\
      3 \have {} & 2
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t] 
      [&6] \\
      6 \have {} & 4 \\
      4 \have {} & 2
    \end{aligned}$
  \end{tabular}}
\frm 2 \hav 1 \end
\frm 1 \hav 0 \end
\end{tabular}\begin{tabular}{@{}p{.25\textwidth}@{}p{.25\textwidth}@{}}
\frm \hav 5 \lor 6 \end
\cases{\begin{tabular}{l|l}
    $\begin{aligned}[t]
      [&5] \\
      \hence {} & 3 \\
      \hence {} & 2
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t] 
      [&6] \\
      \hence {} & 4 \\
      \hence {} & 2
    \end{aligned}$
  \end{tabular}}
\frx 2 \hen 1 \end
\frx \hen 0 \end
\end{tabular}}\]

@@

Our final example is truly diabolical (and sightly unrealistic):
%
\[\hbox{\begin{tabular}{llll}
\raisebox{-\height}{\includegraphics[scale=0.45]{diagrams/diabolical-ref-graph.pdf}}
&
\qquad
\raisebox{-\height}{$\begin{aligned}[t]
\\
\B 0 & \have \B 1 \\
\B 1 & \have \B 2 \\
\B 1 & \have \B 3 \\
\B 2, \B 3 & \have \B 4 \\
\B 2, \B 4 & \have \B 5 \\
\B 3, \B 4 & \have \B 6 \\
\B 2, \B 5, \B 6 & \have \B 7 \\
\B 3, \B 6 & \have \B 8 \\
\B 7, \B 8 & \havebot
\end{aligned}$}
\end{tabular}}\]

We start with

$      |- 7 \lor 8$

Then we identify the nodes that dominate only one branch in the refutation
graph---i.e., that are reachable by navigating backward (upward). Looking at the
graph, we see that $\B 5$ dominates $\B 7$ but not $\bot 8$, and is the only
such node. So we do a case split, and allow ourselves to perform inferences
requiring 5 and 7 in the left branch and 8 in the right branch. First the
right branch:

    [8]
    8 |- 3, 6

Nothing more to do. Any further inferences we would do on it
would need to be repeated in the left branch, which we want to avoid.
Left branch:

         [7]
$    7 |- 2 \lor 5 \lor 6$

    { [2]
    | [5]
$      5 |> 2 \lor 4$
    | [6]}

The 2 and 6 are left alone. We can abbreviate the subcase split to

  $2 \lor 5 \lor 6 |- 2 \lor 4 \lor 6$

In the end, the left branch proves $2 \lor 4 \lor 6$, and
both branches together prove $2 \lor 3 \lor 4 \lor 6$.

This gives us a new case split, but notice that 6 is dominated by every other.
We start with it:

  {[2]
  |[3]
  |[4]
  |[6]
$   6 |- 3 \lor 4$}

i.e. $2 \lor 3 \lor 4$. Since all but one branches are trivial, we abbreviate it
as

$  2 \lor 3 \lor 4 \lor 6 |- 2 \lor 3 \lor 4$

One way to see this is as ``rewriting.'' Similarly with 4:

  {[2]
  |[3]
  |[4]
$   4 |- 2 \lor 3$}

abbreviated as

$  2 \lor 3 \lor 4 |- 2 \lor 3$

i.e. $2 \lor 3$. The rest is reminiscent of our second lasso:

  { [2]
    2 |- 1
  | [3]
    3 |- 1 }
  1 |- 0

Gluing all of this together, we get

@@

Which isn't too bad, considering the spaghetti we started with.

>>>>

\newdimen\bls
\bls=\baselineskip
\newdimen\forhave
\forhave=.8em %%% TYPESETTING

@@

\newcommand\smallbot{\kern-.125em\bot\kern-.125em}
\let\B=\overline
%\newcommand\have{\mathrel{\vdash}}
%\newcommand\have{\mathrel{\;\hookrightarrow\;}}
\newcommand\have{\mathrel{\,\vartriangleright\;}}
\newcommand\hencesym{\blacktriangleright}
\newcommand\hence{\mathrel{\;\hencesym\;}}
%\newcommand\have{\mathrel{\,\rightharpoonup\,}}
\newcommand\havebot{\have\smallbot} %% TYPESETTING
\newcommand\hencebot{\hence\smallbot} %% TYPESETTING

@@

\paragraph{Proof trees.} Proof trees are the standard notation for natural
deduction proofs. For example, the proof graph above is captured by the
following proof tree:
%
\begin{prooftree}
\AXC{$[\B{0}]$}
\UIC{$\B{0}$}
    \AXC{$1$}
        \AXC{$2$}
    \BIC{$1 \land 2$}
        \AXC{$1 \land 2 \longrightarrow 4$}
    \BIC{$4$}
\BIC{$\B{0} \land 4$}
    \AXC{$\kern-4em\B{0} \land 4 \longrightarrow \B{6}$}
\BIC{$\B{6}$}
    \AXC{$[\B{0}]$}
    \UIC{$\B{0}$}
        \AXC{$3$}
            \AXC{$1$}
                \AXC{$2$}
            \BIC{$1 \land 2$}
                \AXC{$1 \land 2 \longrightarrow 4$}
            \BIC{$4$}
        \BIC{$3 \land 4$}
            \AXC{$\kern-3em 3 \land 4 \longrightarrow 5$}
        \BIC{$5$}
    \BIC{$\B{0} \land 5$}
        \AXC{$\kern-5em\B{0} \land 5 \longrightarrow \B{7}$}
    \BIC{$\B{7}$}
\BIC{$\B{6} \land \B{7}$}
    \AXC{$\kern-5em\B{6} \land \B{7} \longrightarrow \bot$}
\BIC{$\bot$}
\end{prooftree}
The proof graph notation is more compact, not least because it enables
sharing of subproofs. For that reason, we will prefer them to proof trees.

\paragraph{Isar proofs.} Isar proofs are a linearization of natural deduction
proofs, but unlike proof trees they enable sharing:
%
\begin{quote}
\keyw{proof}\,~\textit{neg\_clausify} \\
\hbox{}\quad \keyw{assume}\, $\B{0}$ \\
\hbox{}\quad \keyw{have}\, $4$ \keyw{\,by\,} (\textit{metis\, }$1$ $2$) \\
\hbox{}\quad \keyw{have}\, $5$ \keyw{\,by\,} (\textit{metis\, }$3$ $4$) \\
\hbox{}\quad \keyw{have}\, $\B{6}$ \keyw{\,by\,} (\textit{metis\, }$\B{0}$ $4$) \\
\hbox{}\quad \keyw{have}\, $\B{7}$ \keyw{\,by\,} (\textit{metis\, }$\B{0}$ $5$) \\
\hbox{}\quad \keyw{show} \kern.03em$\bot$ \keyw{\kern.05em by\,} (\textit{metis\, }$\B{6}$ $\B{7}$) \\ %% TYPESETTING
\keyw{qed}
\end{quote}
%
The above proof is by contradiction. A direct proof of the above in Isar would
be
%
\begin{quote}
\keyw{proof} -- \\
\hbox{}\quad \keyw{have}\, $4$ \keyw{\,by\,} (\textit{metis\, }$1$ $2$) \\
\hbox{}\quad \keyw{have}\, $5$ \keyw{\,by\,} (\textit{metis\, }$3$ $4$) \\
\hbox{}\quad \keyw{have}\, $6 \lor 7$ \keyw{\,by\,} \textit{metis} \\
\hbox{}\quad $\{$\quad \keyw{assume}\, $6$ \\
\hbox{}\qquad\phantom{$\{$}\keyw{have}\, $0$ \keyw{\,by\,} (\textit{metis\, }$4$ $6$)\quad$\}$ \\
\hbox{}\quad \keyw{moreover} \\
\hbox{}\quad $\{$\quad \keyw{assume}\, $7$ \\
\hbox{}\qquad\phantom{$\{$}\keyw{have}\, $0$ \keyw{\,by\,} (\textit{metis\, }$5$ $7$)\quad$\}$ \\
\hbox{}\quad \keyw{ultimately show}\, $0$ \keyw{\,by\,} (\textit{metis\, }`$6 \lor 7$') \\
\keyw{qed}
\end{quote}
We refer to the Isar tutorial \cite{nipkow-2011-isar} for more information on
the Isar syntax.

@@

We start with a simple proof by contradiction expressed as a proof tree and
using our shorthand notation:
%
\[\hbox{\begin{tabular}{l@{\qquad}l}
\raisebox{-\height}{\includegraphics[scale=0.45]{diagrams/linear-ref-graph.pdf}}
&
\qquad
\raisebox{-\height}{$\begin{aligned}[t]
\\
\B 0, 1 & \have \B 3 \\
2, \B 3 & \have \B 4 \\
1, \B 4 & \havebot
\end{aligned}$}
\end{tabular}}\]
Recall from Sect.\ \ref{sec:proof-notations} that the negated conjecture all all
the proof steps that are tainted by it are shown as negated. Next, we turn the
sequents around using contraposition at the sequent level to avoid all
negations. This gives
\begin{align*}
1, 3 & \have 0 \\
2, 4 & \have 3 \\
1 & \have 4
\end{align*}
Finally, if we reorder the sequents and introduce $\hencesym$ wherever possible,
we obtain a direct proof:

@@

The next two examples are lasso-shaped proofs:
%
\[\hbox{\begin{tabular}{llll}
\raisebox{-\height}{\includegraphics[scale=0.45]{diagrams/lasso-1-ref-graph.pdf}}
&
\qquad
\raisebox{-\height}{$\begin{aligned}[t]
\\
\B 0 & \have \B 1 \\
\B 0 & \have \B 2 \\
\B 1 & \have \B 3 \\
\B 2 & \have \B 4 \\
\B 3, \B 4 & \have \B 5 \\
\B 5 & \have \B 6 \\
\B 6 & \havebot
\end{aligned}$}
&
\qquad\qquad
\raisebox{-\height}{\includegraphics[scale=0.45]{diagrams/lasso-2-ref-graph.pdf}}
&
\qquad
\raisebox{-\height}{$\begin{aligned}[t]
\\
\B 0 & \have \B 1 \\
\B 1 & \have \B 2 \\
\B 2 & \have \B 3 \\
\B 2 & \have \B 4 \\
\B 3 & \have \B 5 \\
\B 4 & \have \B 6 \\
\B 5, \B 6 & \havebot
\end{aligned}$}
\end{tabular}}\]
%
We start with the contradiction proof on the left-hand side. Starting from
$\bot$, it is easy to turn the proof around up to the lasso cycle:
%
\begin{align*}
& \have 6 \\
6 & \have 5 \\
5 & \have 3 \lor 4
\end{align*}
%
When applying the contrapositive to eliminate the negations in $\B 3, \B 4 \have
\B 5$, we obtain a disjunction on the right-hand side of the sequent: $5 \have 3
\lor 4$. To continue from there, we need a case split. Then we can finish then
branches:
%
\[\left[\begin{tabular}{l|l} %% TYPESETTING
    $\begin{aligned}[t]
      [&3] \\
      3 \have {} & 1 \\
      1 \have {} & 0
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t] 
      [&4] \\
      4 \have {} & 2 \\
      2 \have {} & 0
    \end{aligned}$
  \end{tabular}\right]\]

@@

In this case, the key is to observe that it is enough if both branches prove
$2$, and from there we prove the rest. The complete proof is (without and
with $\hencesym$):
\[\hbox{\begin{tabular}{@{}p{.25\textwidth}@{}p{.25\textwidth}@{}}
\frm \hav 5 \lor 6 \end
\cases{\begin{tabular}{l|l}
    $\begin{aligned}[t]
      [&5] \\
      5 \have {} & 3 \\
      3 \have {} & 2
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t] 
      [&6] \\
      6 \have {} & 4 \\
      4 \have {} & 2
    \end{aligned}$
  \end{tabular}}
\frm 2 \hav 1 \end
\frm 1 \hav 0 \end
\end{tabular}\begin{tabular}{@{}p{.25\textwidth}@{}p{.25\textwidth}@{}}
\frm \hav 5 \lor 6 \end
\cases{\begin{tabular}{l|l}
    $\begin{aligned}[t]
      [&5] \\
      \hence {} & 3 \\
      \hence {} & 2
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t] 
      [&6] \\
      \hence {} & 4 \\
      \hence {} & 2
    \end{aligned}$
  \end{tabular}}
\frx 2 \hen 1 \end
\frx \hen 0 \end
\end{tabular}}\]

@@

Our final example is truly diabolical (and sightly unrealistic):
%
\[\hbox{\begin{tabular}{llll}
\raisebox{-\height}{\includegraphics[scale=0.45]{diagrams/diabolical-ref-graph.pdf}}
&
\qquad
\raisebox{-\height}{$\begin{aligned}[t]
\\
\B 0 & \have \B 1 \\
\B 1 & \have \B 2 \\
\B 1 & \have \B 3 \\
\B 2, \B 3 & \have \B 4 \\
\B 2, \B 4 & \have \B 5 \\
\B 3, \B 4 & \have \B 6 \\
\B 2, \B 5, \B 6 & \have \B 7 \\
\B 3, \B 6 & \have \B 8 \\
\B 7, \B 8 & \havebot
\end{aligned}$}
\end{tabular}}\]
%
We start with the contrapositive of the last sequent:
%
\[\have 7 \lor 8\]
%
Then we identify the nodes that dominate only one branch in the refutation graph
(i.e., that are reachable by navigating upward). Looking at the graph, we see
that $\B 5$ dominates $\B 7$ but not $\B 8$, and is the only such node. We
perform a case split and allow ourselves to perform inferences requiring 5 and 7
in the left branch and 8 in the right branch. First the right branch:
%
\begin{align*}
[&8] \\
8 \have {} & 3 \lor 6
\end{align*}
%
There is nothing more we can do for now: Any further inferences we would do in
the right branch would need to be repeated in the left branch. Now we do the
left branch:
%
\begin{align*}
[&7] \\
7 \have {} & 2 \lor 5 \lor 6
\end{align*}
We would now like to perform the inference $5 \have 2 \lor 4$. This requires
a case split:
%
\[\hbox{\begin{tabular}{@{}p{.5\textwidth}@{}p{.5\textwidth}@{}}
\cases{\begin{tabular}{l|r@{}l|l}
[2] & [&5] & [6] \\
    & $5 \have {}$ & $2 \lor 4$
\end{tabular}}\end{tabular}}\]
%
The 2 and 6 subbranches are left alone. Since only one branch of the subcase
split is nontrivial, we find it more aesthetically pleasing to abbreviate it to
%
\[2 \lor 5 \lor 6 \have 2 \lor 4 \lor 6\]
%
Hence, the left branch proves $2 \lor 4 \lor 6$, the right branch proves $3 \lor
6$, and both branches together prove $2 \lor 3 \lor 4 \lor 6$.

This gives us a new case split, but notice that 6 is dominated by 2, 3, and 4.
We start with it:
%
\[\hbox{\begin{tabular}{@{}p{.5\textwidth}@{}p{.5\textwidth}@{}}
\cases{\begin{tabular}{l|l|l|r@{}l}
[2] & [3] & [4] & [&6] \\
& & & $6 \have {}$ & $3 \lor 4$
\end{tabular}}\end{tabular}}\]
This proves $2 \lor 3 \lor 4$. Since all but one branches are trivial, we
abbreviate it:
\[2 \lor 3 \lor 4 \lor 6 \have 2 \lor 3 \lor 4\]
It might help to think of such steps as ``rewriting'' steps, where
6 is rewritten into $3 \lor 4$. Now, 4 is dominated by 2 and 3, so we perform
another case split:
\[\hbox{\begin{tabular}{@{}p{.5\textwidth}@{}p{.5\textwidth}@{}}
\cases{\begin{tabular}{l|l|r@{}l}
[2] & [3] & [&4] \\
& & $4 \have {}$ & $2 \lor 3$
\end{tabular}}\end{tabular}}\]
And we abbreviate it:
\[2 \lor 3 \lor 4 \have 2 \lor 3\]
We are left with the conclusion $2 \lor 3$. The rest is reminiscent of our
second lasso-shaped proof:
\[\hbox{\begin{tabular}{@{}p{.25\textwidth}@{}p{.25\textwidth}@{}}
\cases{\begin{tabular}{l|l}
    $\begin{aligned}[t]
      [&2] \\
      2 \have {} & 1
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t] 
      [&3] \\
      3 \have {} & 1
    \end{aligned}$
  \end{tabular}}
\frm 1 \hav 0 \end
\end{tabular}}\]
%
Gluing all of this together (and using $\hencesym$), we get

@@

\noindent
Which is not too bad, considering the spaghetti we started with.
